\speculake: Hiding Malware in Speculative Execution

### 2018-05-04 paper/architecture.tex
indirect jumps pass

@@

We created a simple trigger program that performs a series of 32 indirect jumps,
using the \texttt{jmpq *\%rax} instruction. Between each jump, we increment
\texttt{\%rax} to account for the number of bytes between jumps. For the last
jump, we load a function pointer into \texttt{\%rax} and do a final indirect
jump using \texttt{callq \%rax}.

In our payload program, we first perform the same 32 indirect jumps. We ensure
the source address of these jumps is the same as in the trigger program by
manually defining their containing function at a fixed address inside a linker
script. We also do the final indirect call to a function pointer, but with two
differences. First, the destination of the function pointer is a different
address, and second, the location of the function pointer in memory is uncached.
This forces the CPU to predict the destination of the indirect call, which it
predicts to be the destination of the function pointer in trigger program. In
the payload program, we purposefully place code at this address which acts as
the \emph{speculative entry point}. Even though the in-order execution of
payload program never executes or even reads from this address, the CPU will
briefly execute instructions there speculatively.

@@

TODO: Note that returns can also be indirect jumps, and rets/jmpq can be mixed
to a certain extent...need to document/quantify how much mixing can be done
still.

>>>>

In \speculake, we cause the CPU to mis-speculate the destination of an indirect
branch in the payload program, causing it to speculatively execute from an
instruction that is never truly executed. We term this speculative destination the
\emph{speculative entry point}.
\speculake uses indirect jumps to allow speculatively executing from \emph{any}
instruction in the payload process' address space. Because it can jump to any
instruction, the malware analyst has a difficult task in determining where a
payload program's speculative entry point is.

In fact, the location this entry point is not determined by the payload
program, but rather the corresponding trigger program. This means that with only
the payload program, an analyst does not posses enough information to find the
speculative entry point.

@@

We observe that Intel CPUs consider three types of x86\_64 indirect branches:
\texttt{retq}, \texttt{callq *\%rax}, and \texttt{jmpq *\%rax}\footnote{other
general purpose registers besides \texttt{\%rax} can be used as well}.
We created a simple trigger program that performs a series of 32 indirect
branches
using \texttt{jmpq *\%rax} instructions. Between each jump, we increment
\texttt{\%rax} to account for the number of bytes between jumps. For the last
jump, we load a function pointer into \texttt{\%rax} and do a final indirect
branch using \texttt{callq *\%rax}.

In our payload program, we first perform the same 32 indirect jumps. We ensure
the source address of these jumps is the same as in the trigger program by
manually defining their containing function at a fixed address inside a linker
script. We also do the final indirect call to a function pointer, but with two
differences. First, the destination of the function pointer is a different
address, and second, the location of the function pointer in memory is uncached.
This forces the CPU to predict the destination of the indirect call while it
waits for the function pointer to load from memory. Due to the similarity
history of branches with the trigger program,
the CPU will (incorrectly) predict the destination to be the same as the one in
the trigger program, which determines the speculative entry point for the
payload. Even though the in-order execution of
payload program never executes or even reads from this address, the CPU will
briefly execute instructions there speculatively.

@@
